CF Nutrition Pre- and Post-CFTR Modulator

### 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
trying to figure out liver labs in python

```{r}
# Import all data
df = read.csv("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Cleaned/analysis.csv")
df$Timepoint = factor(df$Timepoint,levels = c("Pre","Post"))
```

>>>>

```{r}
# Import all data
vit = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx") %>%
  select(`Patient ID`,Date,VitaminA_Retinol:`Vitamin E Gamma`)
fev = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx") %>% 
  rename(`Patient ID` = PatientID) %>%
  select(`Patient ID`,Date,BMI:`FEF25-75 % pred`)
lft = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_LFTs_ThroughDecember2021.xlsx") %>%
  select(`Patient ID`,Date,ALT:GGTP)
# Format dates
vit$Date = as.Date(parse_date(vit$Date,approx = F))
fev$Date = as.Date(parse_date(fev$Date,approx = F))
lft$Date = as.Date(parse_date(lft$Date,approx = F))
# Merge and sort
df = full_join(vit,fev,by = c("Patient ID", "Date"))
df = full_join(df,lft,by = c("Patient ID", "Date"))
df = df %>% arrange(`Patient ID`,Date)
# Conver to numeric
df[,3:ncol(df)] = lapply(df[,3:ncol(df)],as.numeric)
# Format modulator info
mods = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
mods$`Age at Start (years)` = NULL
mods = mods %>% filter(Modulator %in% c("Kalydeco","Trikafta"))
mods$`Earliest Date in Registry` = as.Date(parse_date(mods$`Earliest Date in Registry`,
                                                      approx = T))
mods$`Last Date in Registry` = as.Date(parse_date(mods$`Last Date in Registry`,
                                                      approx = T))
mods = mods %>% pivot_wider(id_cols = `Patient ID`,names_from = Modulator,
                     values_from = `Earliest Date in Registry`)
mods$`Effective Modulator Start` = pmin(mods$Trikafta,mods$Kalydeco,na.rm = T)
mods$fmt = mods$`Effective Modulator Start` == mods$Trikafta
mods$fmk = mods$`Effective Modulator Start` == mods$Kalydeco
mods[,c("fmt","fmk")][is.na(mods[,c("fmt","fmk")])]=0
mods$`First Modulator` = c("Trikafta","Kalydeco")[max.col(mods[,c("fmt","fmk")])]
mods[,c("fmt","fmk")] = NULL
# Put together
df = df %>% filter(`Patient ID` %in% mods$`Patient ID`)
df = left_join(df,mods,by = "Patient ID")
df$`Pre or Post` = factor(df$Date <= df$`Effective Modulator Start`,
                          levels = c("TRUE","FALSE"),labels = c("Pre","Post"))
# Filter and sort again, just in case
df$`Days from Start` = as.numeric(difftime(df$Date,df$`Effective Modulator Start`,units = "days"))
df = df %>% filter(`Days from Start` >= -365*3,
                   `Days from Start` < 0 | `Days from Start` >=90) %>% 
  arrange(`Patient ID`,Date)
# Average pre-modulator values
pre = df %>% group_by(`Patient ID`,`Pre or Post`) %>% summarise_at(vars(BMI:GGTP),mean, na.rm = T) %>% filter(`Pre or Post` == "Pre")
# Find most recent vitamin measure, get any other measures within 30 days
post = df[df$`Pre or Post`=="Post",]
```
